---
title: "p8105_hw1_rl3328"
author: "Ruixi Li"
date: "2023-09-19"
output: html_document
---

Load the moderndive library
```{r, message=FALSE}
library(moderndive)
library(tidyverse)
```

Load the early_january_weather dataset
```{r}
data("early_january_weather")
```
# Problem 1

This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.

## Short descriptive 
```{r}
colnames(early_january_weather)
nrow(early_january_weather)
ncol(early_january_weather)
mean(early_january_weather$temp)
```

* The variables name of this dataset are:`r colnames(early_january_weather)`.
* This data contains `r nrow(early_january_weather)` entries and `r ncol(early_january_weather)` columns.
* The mean temperature is `r mean(early_january_weather$temp)`.

## Scatterplot
we show a scatterplot of `temp` vs `time_hour`. 
```{r scatterplot}
scatter_plot <-  ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()
print(scatter_plot)
ggsave("scatter_plot.pdf", height = 4, width = 6)
```
The temperature in LGA, JFK and EWR seems to increase as time(hour) pass by in January 2013. The temperature decreases when there's rainfall.


# Problem 2
This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

```{r df_create}
factor_levels <- c("Red", "Blue", "Green")

df <- tibble(
  sample = rnorm(10),
  logic = sample > 0,
  char_vector = c("A","B","C","D","E","F","G","H","I","J"),
  factor_vector = factor(sample(factor_levels, 10, replace = TRUE), levels = factor_levels) 
)
print(df)

```


##  Take the mean of each variable in df
```{r mean, eval = FALSE}
mean(pull(df,sample))
mean(pull(df,logic))
mean(pull(df,char_vector))
mean(pull(df,factor_vector))
```
* The mean for variable sample is `r mean(pull(df,sample))`.
* The mean for variable sample is `r mean(pull(df,logic))`.
* mean() function doesn't work on character variables or factor variables.

## Convert variables from one type to another
```{r convert, eval = FALSE}
as.numeric(pull(df,logic))
as.numeric(pull(df,char_vector))
as.numeric(pull(df,factor_vector))
```
* The logical variable and factor variable can be explicitly coverted into numeric, but the character variable cannot.
* It helps partly explain what happens when I try to take the mean.
  * mean() accepts numeric, When I call a function with an argument of the wrong type, R will try to coerce values to a different type so that the function will work. So, the logical variable is converted to numeric(TRUE-1,FALSE-0). That's why mean() works for logical variables.
  * But that doesn't explain why I cannot use mean() to calculate the mean of factor variables.factor variable is categorical variable(either ordinal or nominal), but as.numeric() treats factor variable as ordinal and apply numbers to it, which may be inappropriate.
  
